Group Theory in Physics (Course Note)

### 2016-11-22 Lecture note - Group Theory in Physics/sections/point-group.tex
Adding scanned notes about group "D_n" and "C_n" to group theory notes.
A temporary note for new class - mainly about Wigner- Theorem. Also, some minor modification for my C.V.

Note that since all representations are $1$-dimensional, their
products will keep being irreducible. So having two set of irreducible
representations of $C_n$ and $C_m$, will give us all the irreducible
representations of $C_{nm}$. For example:
% TODO pdf scan of C_6.

\subsection{Dihedral Group \texorpdfstring{$D_n$}{}}
\label{sec:Dihedral-Group}
\begin{defi}[Dihedral Group $D_n$]
\nomenclature{Dihedral Group $D_n$}{\nomrefpage.}
    Dihedral group $D_n$ is the symmetric group of regular $n$-gons
    for $n\geq 3$. For $n=1/2$, it is usually not defined. I found that
    in book \cite{book}, $D_2$ is isomorphic to Klein Group, $D_1$ is
    of course the trivial group.
\end{defi}

@@

% TODO Add scanned notes.

So all in all, $D_n$ can be generated by just two elements. They are:
\begin{fact}
\label{fact:point-group-formal-def-dn}
\begin{align}
    D_n &\equiv \braket{R,S|R^n=1,S^2=1,SRS^{-1}=R^{-1}} \\
        &\equiv \braket{R,S|R^n=1,S^2=1,(SR)^2=1}
\end{align}
Here the first line symbolizes that rotation $R$ and $R^{-1}$ are in
the same conjugacy class. The second is more beautiful and symbolizes
that a roto-reflection is idempotent.
\end{fact}

@@

For $D_3$, it is non-abelian. As discussed previously, it has $3$
conjugacy classes, so it has 3 irreducible representations. Notice
that $D_3/C_3 \cong C_2$, so it inherets the two simple irreducible
representations of $C_2$. This gives two lines in the character
table. Also, $6-1^2-1^2=2^2$, so we have one $2$-dimensional
irreducible representation left.  Next, by fact
\ref{fact:character-of-identity}, we can determine one element in the
character for $2$-dimensional irreducible representation. The rest two
empty blanks can be filled using the orthogonal relations for the
character table.
%TODO Add the scanned example.

@@

>>>>

Note that since all representations are $1$-dimensional, their
products will keep being irreducible. So having two set of irreducible
representations of $C_n$ and $C_m$, will give us all the irreducible
representations of $C_{nm}$. For example:
\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{scans/get-C6-character.pdf}
    \caption{Get $C_6$ Character table}
\end{figure}
\subsection{Dihedral Group \texorpdfstring{$D_n$}{}}
\label{sec:Dihedral-Group}
\begin{defi}[Dihedral Group $D_n$]
\nomenclature{Dihedral Group $D_n$}{\nomrefpage.}
    Dihedral group $D_n$ is the symmetric group of regular $n$-gons
    for $n\geq 3$. For $n=1/2$, it is usually not defined. I found that
    in book \cite{book}, $D_2$ is isomorphic to Klein Group, $D_1$ is
    of course the trivial group.
\end{defi}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{scans/relate-edge-vortex.pdf}
    \caption{relate-edge-Vortex}
\end{figure}
So all in all, $D_n$ can be generated by just two elements. They are:
\begin{fact}
\label{fact:point-group-formal-def-dn}
\begin{align}
    D_n &\equiv \braket{R,S|R^n=1,S^2=1,SRS^{-1}=R^{-1}} \\
        &\equiv \braket{R,S|R^n=1,S^2=1,(SR)^2=1}
\end{align}
Here the first line symbolizes that rotation $R$ and $R^{-1}$ are in
the same conjugacy class. The second is more beautiful and symbolizes
that a roto-reflection is idempotent.
\end{fact}

@@

For $D_3$, it is non-abelian. As discussed previously, it has $3$
conjugacy classes, so it has 3 irreducible representations. Notice
that $D_3/C_3 \cong C_2$, so it inherets the two simple irreducible
representations of $C_2$. This gives two lines in the character
table. Also, $6-1^2-1^2=2^2$, so we have one $2$-dimensional
irreducible representation left.  Next, by fact
\ref{fact:character-of-identity}, we can determine one element in the
character for $2$-dimensional irreducible representation. The rest two
empty blanks can be filled using the orthogonal relations for the
character table.
\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{scans/get-D3-character.pdf}
    \caption{Get $D_3$ Character table}
\end{figure}

@@

\subsection{T-group}

\nomen{$T$ group} is the group of the symmetry of tetrahedrons. It is
a subgroup of $O$, the symmetry of cubic or octahedron (will % TODO ).

A graph is provided in the page 33 of \cite{book}.  Part of this work
has been done in the mid-term exam and is not repreated here.

The conjugacy classes:
$\{E\}$ and $\{3C_2\}$, $\{4C'_2\}$, $\{4C'^2_2\}$.

\paragraph{Character Table}
It has a subgroup of $D_2$, which can be seen from analysing the
number of elments in conjugacy classes. $D_2$ is just $C_3$ discussed
before, and it has three 1D representations. So first three line of
the character is simple to get.

The remaing representation is $3$-dimensional, so we can again fill
one element of the character table. The rest can be easily filled by
the orthogonal relations.

Details on page 66 and 33 of \cite{book}.
